Negative Selection in the 20th Century

### 2020-05-15 why-negative-selection.Rmd
Fix paths in _drake.R; split time regressions by own YOB always. Put data files in gitignore. Stop storing drake history.

* Divide siblings regressions by own YOB, into early and late.
  - Use this to confirm the basic "selection decreasing" story

* Plot average scores over time, as an intro graph
  - Can we work out how to plot selection effects on the "same scale"?
  - Let X_i be i's polygenic score.
  - If E(X_parents) = 1/n \sum X_i in the parent's generation, then
    E(X_kids) = 1/n \sum X_i K_i, in the children's gernation, where K_i is 
    the number of kids i has. 
  - Up to a change of means, this is the same as the sample correlation between
    X and K. 
  - Do we calculate correlations per year, then plot? Probably too low N....
    Maybe calculate the correlation over the whole sample but then use
    it with each year's X_i profile

* How much does age at first live birth "account for" selection?

@@

summary_period <- res_period %>% 
      select(-std.error, -statistic) %>% 
      pivot_wider(
        names_from  = period, 
        values_from = c(estimate, conf.low, conf.high, p.value)
      ) %>% 
      group_by(score_name, children) %>% 
      mutate(
        significant = diff.p.value < 0.05/n_tests,
        increasing  = estimate_early < estimate_late,
        size_inc    = abs(estimate_early) < abs(estimate_late),
        change_sign = sign(estimate_early) != sign(estimate_late),
        Change      = case_when(
                        ! significant ~ "Insignificant",
                        change_sign  & increasing   ~ "Change sign to +",
                        change_sign  & ! increasing ~ "Change sign to -",
                        increasing   & size_inc   ~ "Increasing +",
                        ! increasing & size_inc   ~ "Decreasing -",
                        increasing   & ! size_inc ~ "Increasing towards 0",
                        ! increasing & ! size_inc ~ "Decreasing towards 0"
                      )
      ) %>% 
      select(score_name, children, Change, increasing, significant) %>% 
      mutate(children = ifelse(children, "Respondents", "Parents")) %>% 
      pivot_wider(names_from = children, 
        values_from = c(Change, increasing, significant)) %>% 
      mutate(overall = case_when(
              increasing_Respondents   & increasing_Parents   ~ 
                "Consistent increase",
              ! increasing_Respondents & ! increasing_Parents ~ 
                "Consistent decrease",
              increasing_Respondents & ! increasing_Parents
                 & (significant_Parents | significant_Respondents) ~ 
                "Changes direction",
              TRUE ~ 
              "Other"
            ))

>>>>

* Divide siblings regressions by own YOB, into early and late.
  - Use this to confirm the basic "selection decreasing" story
  - Actually almost no change.

* Plot average scores over time, as an intro graph. DONE
  - Can we work out how to plot selection effects on the "same scale"?
  - Simplest thing would be to use birth dates

* How much does age at first live birth "account for" selection?
  - Need f.3872 age at flb for women with one child

@@

summary_period <- res_period %>% 
      select(-std.error, -statistic) %>% 
      pivot_wider(
        names_from  = period, 
        values_from = c(estimate, conf.low, conf.high, p.value)
      ) %>% 
      group_by(score_name, children) %>% 
      mutate(
        significant = diff.p.value < 0.05/n_tests,
        sign_early  = sign(estimate_early),
        sign_late   = sign(estimate_late),
        size_inc    = abs(estimate_early) < abs(estimate_late),
        change_sign = sign_early != sign_late,
        Change      = case_when(
                        ! significant ~ "Insignificant",
                        change_sign   ~ "Change sign",
                        size_inc      ~ "Size increasing",
                        ! size_inc    ~ "Size decreasing"
                      )
      ) %>% 
      select(score_name, children, Change, significant, change_sign, 
        sign_early, sign_late) %>% 
      mutate(children = ifelse(children, "Respondents", "Parents")) %>% 
      pivot_wider(names_from = children, 
        values_from = c(Change, significant, change_sign, sign_early, 
        sign_late)) %>% 
      mutate(overall = case_when(
              ! change_sign_Parents & ! change_sign_Respondents & 
                sign_late_Parents == sign_early_Parents ~ 
                "Consistent",
              TRUE ~ "Changes direction"
            ))
